---
title: crosswalkr
output: md_document
---

```{r, include = FALSE}
options(width = 100)
library(dplyr)
```

![](https://img.shields.io/badge/dev-beta-red.svg)
[![GitHub release](https://img.shields.io/github/release/btskinner/crosswalkr.svg)](https://github.com/btskinner/crosswalkr)
[![Travis-CI Build Status](https://travis-ci.org/btskinner/crosswalkr.svg?branch=master)](https://travis-ci.org/btskinner/crosswalkr)


This package offers a pair of functions, `renamefrom()` and
 `encodefrom()`, for renaming and encoding data frames using external
 crosswalk files. It is especially useful when constructing master
 data sets from multiple smaller data sets that do not name or encode
 variables consistently across files. Based on `renamefrom` and
 `encodefrom`
 [Stata commands written by Sally Hudson and team](https://github.com/slhudson/rename-and-encode).

### Install

Install the latest development version from Github with

```{r, eval = FALSE}
devtools::install_github('btskinner/crosswalkr')
```

### Dependencies

This package relies on the following packages, available in CRAN:

* haven
* labelled
* readr
* readxl
* tibble

## Available functions

### `renamefrom()`

#### Examples

```{r}
library(crosswalkr)

## starting data frame
df <- data.frame(state = c('Kentucky','Tennessee','Virginia'),
                 fips = c(21,47,51),
                 region = c('South','South','South'))

## crosswalk with to convert old names to new names with labels
cw <- data.frame(old_name = c('state','fips'),
                 new_name = c('stname','stfips'),
                 label = c('Full state name', 'FIPS code'))

## convert old to new using labels in crosswalk file
df1 <- renamefrom(df, cw, old_name, new_name, label)
df1

## convert old to new using old names as labels
df2 <- renamefrom(df, cw, old_name, new_name, name_label = TRUE)
df2

## convert old to new, but keep unmatched old names in data frame
df3 <- renamefrom(df, cw, old_name, new_name, drop_extra = FALSE)
df3
```

### `encodefrom()`

#### Examples

```{r}

## starting data frame
df <- data.frame(state = c('Kentucky','Tennessee','Virginia'),
                 stfips = c(21,47,51),
                 cenregnm = c('South','South','South'))

## use state crosswalk data file from package
cw <- get(data(stcrosswalk))

## create new column with factor-encoded values
df$state2 <- encodefrom(df, state, cw, stname, stfips, stabbr)
df

## create new column with labelled values
df_tbl$state2 <- encodefrom(df_tbl, state, cw, stname, stfips, stabbr)

## create new column with factor-encoded values (ignores the fact that df_tbl is a tibble)
df_tbl$state3 <- encodefrom(df_tbl, state, cw, stname, stfips, stabbr, ignore_tibble = TRUE)

## show results
haven::as_factor(df_tbl)
haven::zap_labels(df_tbl)
```
